Global Assessments of Fishery Status need Better Data more than Better Models

### 2020-07-30 documents/ovando-etal-assessing-global-fisheries.Rmd
off to github

# Introduction (500 words)

@@

# Results  (2000 words)

@@

## Case Studies

Two case studies provide examples of how augmenting catch data alone with local and globally available data can provide dramatically different estimates of local stock status. For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region. We can then compare the estimates of stock status that come from using the catch-data alone (with a heuristic algorithm in the manner of @martell2013), with the catch data along with priors on terminal depletion informed by the swept-area-ratio data, and using the swept-area ratio data together with a local abundance index. By it's nature the catch-only heuristic produces a bi-model distribution, with most fisheries reported to have a B/B~MSY~ of 1.5 and an F/F~MSY~ less than 1 (Fig.\@ref(fig:east-india)-heuristic). Local experts put the swept-area-ratio (SAR) value for the region at an incredibly high 17. Using this value to inform a prior on terminal stock status dramatically changes our picture of stock status in the Eastern Bay of Bengal, with now all stocks being overfished and experiencing high levels of overfishing (Fig.\@ref(fig:east-india)-sar).

<!-- Evaluating the validity of data-limited assessments methods can be challenging. Simulation testing provides one avenue, but it can be difficult to create a simulation environment that adequately captures the problems of real-world ecosystems and data. We can also compare results between data-limited assessment and more data (or resource) rich gold-standard stock assessments. However, formally stock assessed fisheries represent a very biased subset of global fisheries, and it is entirely possible that a method that performs well for the kinds of species and fisheries that are commonly stock assessed might struggle in the kinds of fisheries that typically are unassessed.  -->

<!-- Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region. As outlined in Table.\@ref(tab:status-tab), while we often have no proof that results finding that for example fisheries in south-east Asia have similar stock status to those in Alaska, very few experts would agree that this is a plausible result XX tough statement to make in a paper xx. We use two case studies here to present the potential of additional, broadly applicable, data to potentially improve on "catch-only" methods.  -->

If all the stocks in question were demersal, shared similar life histories, and were all caught in the same multi-species fleet, this result might make sense. However, imposing the same prior on terminal stock status for all fisheries in a region is not ideal. Supplementing these data with a local index of abundance provides a more nuanced result, with most stocks estimated as having B/B~MSY~ less than 1, but with a much wider range of fishing mortality rates reflecting difference in recent catch histories among individual stocks (Fig.\@ref(fig:east-india)-"cpue + sar"). While we do not know the true state of stocks in the Eastern Bay of Bengal, inclusion of local data provide a picture more in line with local perception XX.

```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio. Each point is a stock."}
ei_kobe_plot 
```

It should be no surprise that incorporating locally tailored data produces more locally plausible results. Global assessment at the moment though depend on few broadly available datasets. We used `sraplus` to assess the status of stocks in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from [@rousseau2019] (xx technically this is the updated dataset xx). Catch histories for Europe show a steady increase reaching a peak in the 1990s, and a decline and stabilization since. Current catches in Europe are much higher than those at the start of the catch history. Catches in northeast Asia have steadily increased over time, with some stabilization as of late. For both regions, recent catches expressed as a fraction of maximum catch, they key metric from the perspective of a catch-only heuristic, are both high. However, the two regions exhibit markedly different effort histories according to @rousseau2019, with effort in Europe peaking with catches and subsequently declining, while effort in northeast Asia has increased steadily over time. As a result, using a "catch-only" heuristic, both Europe and Northeast Asia are estimated to have very similar stock status. Adding in an index of abundance, calculated for each stock as catch divided by the effective harvest rate based on the effort data for the region, produces a very different picture of stock status and introduces some contrast between the two regions. Based on the catch and effort data, `sraplus` estimates that more stocks in Northeast Asia are overfished / experiencing overfishing than in Europe. However, the shared effort series across these stocks also produces a large number of effectively unfished stocks; a highly unlikely result for both regions (Fig.\@ref(fig:eu-v-na)).

@@

These case studies reveals the potential and challenges of using broadly available data for global fishery assessment. On the one hand, for both case studies moving beyond catch data alone produced a more plausible set of results in some ways. However, both case studies also raise a host of new issues. How realistic is it to share one index of abundance and SAR value for the entire eastern Bay of Bengal? Similarly, the sharing of a common effort series at the continental level does provide increased separation between European and northeast Asian stocks that may be plausible, but also estimates a suspicious amount of stocks to be essentially unfished.

Extending these issues to the global level, we  used a variety of data sources to estimate stock status for fisheries listed in RAM. Our goal is to simulate a macro-level global assessment, where data are available at regional levels, but not for specific fisheries. As a proof of concept, we first fit models to abundance indices from each stock in RAM. There are two main reasons why an assessment model might produce poor results: model misspecification and poor data. Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries. Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`. Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality. Overall the `sraplus` estimates using the RAM data are reasonably good. The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias. The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time. Looking across space, we see clear geographic patterns in model performance. The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada). The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.

Performance limitations then are likely to arise less from model misspecification but from the quality of the data themselves. These becomes even clearer once we consider the performance of `sraplus` models fit to globally available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4. The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1. This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])`.

@@

# Discussion (500 words)

@@

Our estimates of global stock status vary wildly depending on which source of data we use. Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information? Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)). Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used. These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability. The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations. Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks. None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RAM (XX RAM matches? need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).

@@

As we should expect having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 20%. In terms of data however, information on fishing mortality rates (F/F~MSY~) in the most recent year was almost as useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot). Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance. This is because we varied both the length and quality of the index of abundance, and on average an even noisy estimate of recent F/F~MSY~ may be more valuable than a short and noisy abundance index.

@@

# Methods (3000 words)

>>>>

# Introduction

@@

# Results

@@

<!-- ## Case Studies  -->

<!-- Two case studies provide examples of how augmenting catch data alone with local and globally available data can provide dramatically different estimates of local stock status. For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region. We can then compare the estimates of stock status that come from using the catch-data alone (with a heuristic algorithm in the manner of @martell2013), with the catch data along with priors on terminal depletion informed by the swept-area-ratio data, and using the swept-area ratio data together with a local abundance index. By it's nature the catch-only heuristic produces a bi-model distribution, with most fisheries reported to have a B/B~MSY~ of 1.5 and an F/F~MSY~ less than 1 (Fig.\@ref(fig:east-india)-heuristic). Local experts put the swept-area-ratio (SAR) value for the region at an incredibly high 17. Using this value to inform a prior on terminal stock status dramatically changes our picture of stock status in the Eastern Bay of Bengal, with now all stocks being overfished and experiencing high levels of overfishing (Fig.\@ref(fig:east-india)-sar).  -->

<!-- <!-- Evaluating the validity of data-limited assessments methods can be challenging. Simulation testing provides one avenue, but it can be difficult to create a simulation environment that adequately captures the problems of real-world ecosystems and data. We can also compare results between data-limited assessment and more data (or resource) rich gold-standard stock assessments. However, formally stock assessed fisheries represent a very biased subset of global fisheries, and it is entirely possible that a method that performs well for the kinds of species and fisheries that are commonly stock assessed might struggle in the kinds of fisheries that typically are unassessed.  --> -->

<!-- <!-- Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region. As outlined in Table.\@ref(tab:status-tab), while we often have no proof that results finding that for example fisheries in south-east Asia have similar stock status to those in Alaska, very few experts would agree that this is a plausible result XX tough statement to make in a paper xx. We use two case studies here to present the potential of additional, broadly applicable, data to potentially improve on "catch-only" methods.  --> -->

<!-- If all the stocks in question were demersal, shared similar life histories, and were all caught in the same multi-species fleet, this result might make sense. However, imposing the same prior on terminal stock status for all fisheries in a region is not ideal. Supplementing these data with a local index of abundance provides a more nuanced result, with most stocks estimated as having B/B~MSY~ less than 1, but with a much wider range of fishing mortality rates reflecting difference in recent catch histories among individual stocks (Fig.\@ref(fig:east-india)-"cpue + sar"). While we do not know the true state of stocks in the Eastern Bay of Bengal, inclusion of local data provide a picture more in line with local perception XX.  -->

<!-- ```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio. Each point is a stock."} -->
<!-- ei_kobe_plot  -->
<!-- ``` -->

We used `sraplus` to assess the status of stocks in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from [@rousseau2019] (xx technically this is the updated dataset xx). Catch histories for Europe show a steady increase reaching a peak in the 1990s, and a decline and stabilization since. Current catches in Europe are much higher than those at the start of the catch history. Catches in northeast Asia have steadily increased over time, with some stabilization as of late. For both regions, recent catches expressed as a fraction of maximum catch, they key metric from the perspective of a catch-only heuristic, are both high. However, the two regions exhibit markedly different effort histories according to @rousseau2019, with effort in Europe peaking with catches and subsequently declining, while effort in northeast Asia has increased steadily over time. As a result, using a "catch-only" heuristic, both Europe and Northeast Asia are estimated to have very similar stock status. Adding in an index of abundance, calculated for each stock as catch divided by the effective harvest rate based on the effort data for the region, produces a very different picture of stock status and introduces some contrast between the two regions. Based on the catch and effort data, `sraplus` estimates that more stocks in Northeast Asia are overfished / experiencing overfishing than in Europe. However, the shared effort series across these stocks also produces a large number of effectively unfished stocks; a highly unlikely result for both regions (Fig.\@ref(fig:eu-v-na)).

@@

Extending this idea to the global level, we  used a variety of data sources to estimate stock status for fisheries listed in RAM. Our goal is to simulate a macro-level global assessment, where data are available at regional levels, but not for specific fisheries. As a proof of concept, we first fit models to abundance indices from each stock in RAM. We assess performance by three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy is calcualted as the proportion of times that a given dataset results in a stock falling into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished). There are two main reasons why an assessment model might produce poor results: model misspecification and poor data. Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries. Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`. Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality. Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(teb:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.

<!-- The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias. The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time. Looking across space, we see clear geographic patterns in model performance. The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada). The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.  -->

Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves. These becomes even clearer once we consider the performance of `sraplus` models fit to globally available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4. The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1. This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])`.

@@

# Discussion

@@

<!-- Our estimates of global stock status vary wildly depending on which source of data we use. Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information? Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)). Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used. These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability. The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations. Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks. None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RAM (XX RAM matches? need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).  -->

@@

Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 20%. In terms of data however, information on fishing mortality rates (F/F~MSY~) in the most recent year was almost as useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot). Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance. This is because we varied both the length and quality of the index of abundance, and on average an even noisy estimate of recent F/F~MSY~ may be more valuable than a short and noisy abundance index.

@@

# Methods
